---
title: "Ecological Analysis with R: Night Session #1"
output: html_document
incremental: true
always_allow_html: yes

--- ### For tonight's class we will be going through the same workflow
that we went through before, but we will be answering a real
ecological question, and we will be going into some more advanced
topics.  While we all be working with the same data, there will be
options with the analysis for anyone looking for a bit of a challenge.

## Once again remember our workflow.
0. Get a question and decide on analyses
1. Get the data
2. Check the data
3. Prep the data
4. Visualize patterns
5. Statistical Testing
6. Export results


###0.) In this case, the question is decided: __Are there differences
   in productivity between trees in Washington and trees in
   Massachusetts?__

###1.) Download the data from GitHub https://github.com/HarvardForest/Rworkshop/tree/master/data
    + Setup your project folder 
    + Download CleanedMATrees.csv & CleanedWATrees.csv
	+ If you would like to start out with a bit of a challenge also download: MA_PLOT.csv & WA_PLOT.csv
	+ DO NOT DOUBLE CLICK OR OPEN WITH EXCEL
	+ Read it into R 


###2.)  Look at the data.  What columns will be useful for us to
   answer the question? Are there any pieces that we don't know what
   they are? Make sure to compare the states!

__Challenge Mode:__ Look at the plot data, see if you can figure out
the connections between plot and tree. Figure out how many unique
trees there are and how many unique plots.


###3.)   We want to prep the data for our analysis. Seems easy, let's just sum all the carbon in each state! 



``` {r,eval = F}

install.packages(plyr)
ddply(DATASET,.(COLUMNNAME),summarise,TotalCarbon=sum(CARBON_AG))

```

#####What are the issues with this?
  + Different amounts of area in state.
  + Different amounts of forested land.
  + How to correct for these?

__Challenge Mode:__
Extract only the plots that were sampled and are forested (```PLOT_STATUS_CD == 1```)
Link the PLOT table and the TREE table using ```merge()``` or ```join()```.


###4.) Plotting using ggplot

The ggplot approach to plotting produces visually appealing graphics
fairly easily and is currently one of the more commonly used graphics
packages in R.

```{r , eval=F}

ggplot(data, aes(x=x_variable, y=y_variable)) +
    geom_point(shape=1)

```

###4.) Plotting using ggplot

```{r , eval=F}

library(ggplot2)
dat <- rbind(MAtrees,WAtrees)
dat$state <- c(rep('MA',nrow(MAtrees)),rep('WA',nrow(WAtrees)))

ggplot(dat, aes(x=CN, y=DIA, color=state)) +
    geom_point(shape=1) +
    scale_colour_hue(l=50) +
    geom_smooth(method=lm, se=TRUE)

```

###4.) Plotting using ggplot

```{r , eval=F}

our.plot <- ggplot(dat, aes(x=CN, y=DIA, color=state))

our.plot +  geom_point(shape=1) +
    scale_colour_hue(l=50) +
    geom_smooth(method=lm, se=TRUE)


```


###5.) Statistics


```{r , eval=F}

dia.WA14 <- WAtrees[WAtrees[,'INVYR'] == 2014,'DIA']
dia.MA14 <- MAtrees[MAtrees[,'INVYR'] == 2014,'DIA']

t.test(dia.WA14,dia.MA14)

```
###5.) Statistics

```{r , eval=F}

fit <- aov(DIA~state,data=dat)
summary(fit)

capture.output(summary(fit))
dia.anova <- capture.output(summary(fit))
write.csv(dia.anova)


```

###5.) Statistics - Assumptions

```{r , eval=F}

ggplot(dat, aes(x=DIA,color=state)) + 
    geom_histogram()

```
